Unifying Lower Bounds on Prediction Dimension of Consistent Convex Surrogates

### 2021-05-17 cvx-flats-neurips21/cvx-flats.tex
Adding Tikz figures for diagram

\documentclass{article} % Anonymized submission
% \renewcommand{\includegraphics}[1][1]{}

\usepackage{neurips_2021}

@@

\PassOptionsToPackage{numbers, sort, compress}{natbib}

\usepackage{mathtools, verbatim}
%\usepackage[thmmarks, thref, amsthm]{ntheorem}
\usepackage{color}
\definecolor{darkblue}{rgb}{0.0,0.0,0.2}
\hypersetup{colorlinks,breaklinks,
	linkcolor=darkblue,urlcolor=darkblue,
	anchorcolor=darkblue,citecolor=darkblue}
\usepackage{wrapfig}
\usepackage[small]{caption}
\usepackage[colorinlistoftodos,textsize=tiny]{todonotes} % need xargs for below
%\usepackage{accents}
\usepackage{bbm}
\usepackage{xspace}
\usetikzlibrary{arrows}
\usetikzlibrary{shapes}

@@

@@

Given a target loss $\ell$, we can define a statistic $\gamma$, the property it elicits.
Intuitively, consistency of a surrogate $L$ with respect to $\ell$ and $\gamma$ are equivalent, i.e. in both cases estimates should converge to values that minimize $\ell$-loss.
We formalize this by letting $\propdis$ be the $\ell$-regret, yielding Lemma~\ref{lem:consistent-loss-implies-prop}, proven in Appendix~\ref{app:misc-omitted-proofs}.
%If $\gamma$ is a property elicited by a target loss $\ell$, then taking $\propdis$ to be the $\ell$-regret makes the two definitions are equivalent, as stated in Lemma~\ref{lem:consistent-loss-implies-prop}, proven in Appendix~\ref{app:misc-omitted-proofs}.% for the appropriate choice of $\propdis$.

@@

\section{Discrete-valued predictions}\label{sec:finite-calib}

\begin{figure}[htb]
	\centering
	\begin{tabular}{cc}
%		\includegraphics[width=.35\textwidth]{tikz/fsd-bound.pdf} &
%		\includegraphics[width=.35\textwidth]{tikz/flats-bound.pdf} \\
	\end{tabular}
	\caption{(Left) Feasible subspace dimension $\Sc_{\gamma_\bot}(\bullet) = 2$ and $\Sc_{\gamma_\bot}({\color{blue}\star}) = 1$, giving the bound $\conscvx(\ell^{abs}) \geq 3- 1-1 = 1$.
          (Right) No $1$-flat through $\bullet$ (a line since $\bullet \in \relint{\simplex})$  stays fully contained in $\gamma_\bot$, so $\conscvx(\ell^{abs}) \geq 2$.}
	\label{fig:fsd-flats-abstain}
\end{figure}

\section{Continuous-valued predictions}\label{sec:contin-consis}

@@

%\bibliographystyle{plainnat}
\bibliography{diss,extra}

@@

>>>>

\documentclass[11pt]{article} % Anonymized submission
% \renewcommand{\includegraphics}[1][1]{}

\usepackage[nonatbib]{neurips_2021}

@@

\usepackage[numbers, sort, compress]{natbib}

\usepackage{mathtools, verbatim}
%\usepackage[thmmarks, thref, amsthm]{ntheorem}
%\usepackage{color}
\usepackage[dvipsnames]{xcolor}
\definecolor{darkblue}{rgb}{0.0,0.0,0.2}
\hypersetup{colorlinks,breaklinks,
	linkcolor=darkblue,urlcolor=darkblue,
	anchorcolor=darkblue,citecolor=darkblue}
\definecolor{myblue}{rgb}{0,0.3,0.8} \definecolor{myorange}{rgb}{0.9,0.5,0}
\usepackage{wrapfig}
\usepackage[small]{caption}
\usepackage[colorinlistoftodos,textsize=tiny]{todonotes} % need xargs for below
%\usepackage{accents}
\usepackage{bbm}
\usepackage{xspace}

\usepackage{tikz,pgfplots,tikz-3dplot}
\usetikzlibrary{arrows.meta,shapes.misc,positioning,fit,calc,intersections,through,decorations.markings}
\tdplotsetmaincoords{55}{135}        % for simplex diagrams

\tikzstyle{cell}=[dashed,thick]
\tikzstyle{simplex}=[thick]
\newcommand{\tikzfigscale}{2.2}
\newcommand{\placefiglabel}[1]{\node at (1.1,0,1.7) {#1};}
\usetikzlibrary{arrows}
\usetikzlibrary{shapes}

@@

\jessie{If, instead of a target loss $\ell$, we have a target property as in Quadrants 2 and 4, we might want to equivalently define consistency with respect to a property.  
Intuitively, this definition is similar, but is equipped with a metric-like function to measure equivalence to a target property value.  
See Appendix~\ref{} for our precise definition, which we show is equivalent to consistency with respect to target loss.
This allows us to ``translate'' results about target losses to be about the properties these losses elicit, and therefore apply our results to all four Quadrants.}\jessiet{Suggestion: replace following text up to Lemma 1 with this.}

@@

Given a target loss $\ell$, we can define a statistic $\gamma$, the property it elicits.
Intuitively, consistency of a surrogate $L$ with respect to $\ell$ and $\gamma$ are equivalent, i.e. in both cases estimates should converge to values that minimize $\ell$-loss.
We formalize this by letting $\propdis$ be the $\ell$-regret, yielding Lemma~\ref{lem:consistent-loss-implies-prop}, proven in Appendix~\ref{app:misc-omitted-proofs}.\jessiet{Move this proof if property consistency is moving to its own appendix as well.}
%If $\gamma$ is a property elicited by a target loss $\ell$, then taking $\propdis$ to be the $\ell$-regret makes the two definitions are equivalent, as stated in Lemma~\ref{lem:consistent-loss-implies-prop}, proven in Appendix~\ref{app:misc-omitted-proofs}.% for the appropriate choice of $\propdis$.

@@

\section{Illustration: geometric intuition for the four quadrants}
We can apply Lemma~\ref{lem:convex-flats-inf-dim} to construct lower bounds on convex consistency dimension for targets across all four quadrants.
This is demonstrated in Figure~\ref{tab:illustrate-quadrant-apps}.

\begin{figure}
  \begin{tikzpicture} [scale=\tikzfigscale, thick, tdplot_main_coords]
% standard tikz coordinate definition using x, y, z coords
\coordinate (orig) at (0,0,0);

\pgfmathsetmacro{\ph}{0.54}
\pgfmathsetmacro{\pm}{0.23}
\pgfmathsetmacro{\pl}{0.23}
\coordinate (p) at (\ph,\pm,\pl);
\coordinate[label=below:${y=-1\!\!\!\!}$] (h) at (1,0,0);
\coordinate[label=below:${y=1}$] (m) at (0,1,0);
\coordinate[label=above:${y=0}$] (l) at (0,0,1);

% draw axes
\draw[simplex] (h) -- (m) -- (l) -- (h);

\begin{scope}
\clip (h) -- (m) -- (l) -- (h);
\foreach \c in {0,0.05,...,1}
{
	\pgfmathsetmacro\aa{\c}
	\coordinate (A) at ($(\aa,1-\aa,0)$);
	\coordinate (B) at ($(\aa,1-\aa,2)$);
	\draw[myblue,thin] (A) -- (B);
}
\draw[orange, very thick,dashed] (0.3,0.7,0) -- (0.3,0.7,2);
%\draw[orange, very thick] (0.7,0.3,0) -- (0.7,0.3,2);
\end{scope}
\node[inner sep=2pt] (levset1) at (0.3,0.7,-.3) {$\Gamma_{0.4}\!\!$};
%\node[inner sep=2pt] (levset2) at (0.7,0.3,-.3) {$\Gamma_{\!\!-0.4}\!\!\!\!$};
%\draw[-latex,thin] (levset2) edge[in=270,out=90] (0.7,0.3,0);
\draw[-latex,thin] (levset1) edge[in=270,out=90] (0.3,0.7,0);
\placefiglabel{(Q3)}
\end{tikzpicture}
\hfill
\begin{tikzpicture} [scale=\tikzfigscale, thick, tdplot_main_coords]
% standard tikz coordinate definition using x, y, z coords
\coordinate (orig) at (0,0,0);

\pgfmathsetmacro{\ph}{0.54}
\pgfmathsetmacro{\pm}{0.23}
\pgfmathsetmacro{\pl}{0.23}
\coordinate (p) at (\ph,\pm,\pl);
\coordinate[label=below:${y=-1\!\!\!\!}$] (h) at (1,0,0);
\coordinate[label=below:${y=1}$] (m) at (0,1,0);
\coordinate[label=above:${y=0}$] (l) at (0,0,1);

% draw axes
\draw[simplex] (h) -- (m) -- (l) -- (h);

\begin{scope}
\clip (h) -- (m) -- (l) -- (h);
\foreach \c in {0,0.1,...,1}
{
	\pgfmathsetmacro\aa{0.5*(sqrt(1-\c)+1)}
	\pgfmathsetmacro\bb{0.5*\c}
	\coordinate (A) at ($(\aa,1-\aa,0)$);
	\coordinate (B) at ($(1-\aa,\aa,0)$);
	\coordinate (C) at ($(\bb,\bb,1-2*\bb)$);
	\draw[myblue,thin] (C) parabola (A);
	\draw[myblue,thin] (C) parabola (B);
}
\def\c{0.8}
\pgfmathsetmacro\aa{0.5*(sqrt(1-\c)+1)}
\pgfmathsetmacro\bb{0.5*\c}
\coordinate (A) at ($(\aa,1-\aa,0)$);
\coordinate (B) at ($(1-\aa,\aa,0)$);
\coordinate (C) at ($(\bb,\bb,1-2*\bb)$);
\draw[myorange, very thick] (C) parabola (A);
\draw[myorange, very thick] (C) parabola (B);
\draw[myorange, dashed] (A) -- (0,1/3,2/3);
\end{scope}
\def\c{0.8}
\pgfmathsetmacro\aa{0.5*(sqrt(1-\c)+1)}
\pgfmathsetmacro\bb{0.5*\c}
\node (levset) at (0.5,0.5,-.3) {$\Gamma_{0.8}$};
\draw[-latex,thin] (levset) edge[in=270,out=180] ($(\aa,1-\aa,0)$);
\draw[-latex,thin] (levset) edge[in=270,out=0] ($(1-\aa,\aa,0)$);
\placefiglabel{(Q4)}
\end{tikzpicture}
\hfill
\begin{tikzpicture} [scale=\tikzfigscale, thick, tdplot_main_coords]
% standard tikz coordinate definition using x, y, z coords
\coordinate (orig) at (0,0,0);

\pgfmathsetmacro{\ph}{0.54}
\pgfmathsetmacro{\pm}{0.23}
\pgfmathsetmacro{\pl}{0.23}
\coordinate (p) at (\ph,\pm,\pl);
\coordinate (uniform) at (1/3,1/3,1/3);
\coordinate[label=below:${y=-1\!\!\!\!}$] (h) at (1,0,0);
\coordinate[label=below:${y=1}$] (m) at (0,1,0);
\coordinate[label=above:${y=0}$] (l) at (0,0,1);

% draw axes
\draw[simplex] (h) -- (m) -- (l) -- (h);

\begin{scope}
\clip (h) -- (m) -- (l) -- (h);
\draw[myblue] (1/2,1/2,0) -- (1/2, 0, 1/2);
\draw[myblue] (0,1/2,1/2) -- (1/2, 0, 1/2);
\draw[myblue] (0,1/2,1/2) -- (1/2,1/2, 0);
\draw[myorange, dashed] (1/2,0,1/2) -- (0,1,0);
\draw[myorange, dashed] (2/3,1/3,0) -- (0,1/3,2/3);
\node (uniformdot) at (1/3,1/3,1/3) {$\bullet$};
%\draw[orange, very thick] (0.7,0.3,0) -- (0.7,0.3,2);
\end{scope}
\node[inner sep=2pt] (levset1) at (3/4,1/8,1/8) {$\Gamma_{1}$};
\node[inner sep=2pt] (levsetabs) at (5/12,5/12,1/6) {$\Gamma_{\bot}$};
%\draw[-latex,thin] (levset2) edge[in=270,out=90] (0.7,0.3,0);
%\draw[-latex,thin] (levset1) edge[in=270,out=90] (0.3,0.7,0);
\placefiglabel{(Q1)}
\end{tikzpicture}
\hfill
\begin{tikzpicture} [scale=\tikzfigscale, thick, tdplot_main_coords]
% standard tikz coordinate definition using x, y, z coords
\coordinate (orig) at (0,0,0);

\pgfmathsetmacro{\ph}{0.54}
\pgfmathsetmacro{\pm}{0.23}
\pgfmathsetmacro{\pl}{0.23}
\coordinate (p) at (\ph,\pm,\pl);
\coordinate (uniform) at (1/3,1/3,1/3);
\coordinate[label=below:${y=-1\!\!\!\!}$] (h) at (1,0,0);
\coordinate[label=below:${y=1}$] (m) at (0,1,0);
\coordinate[label=above:${y=0}$] (l) at (0,0,1);

% draw axes
\draw[simplex] (h) -- (m) -- (l) -- (h);

\begin{scope}
\clip (h) -- (m) -- (l) -- (h);
\draw[myblue,thin] (1/4,0,3/4) -- (0,1/4, 3/4);
\draw[myblue,thin] (1/8,1/8,3/4) -- (1/2, 1/2,0);
\def\e{1/16};
\draw[orange, dashed] (1/8 + \e, 0, 1 - 1/8 - \e) -- (0, 1 - 1/8 - \e, 1/8 + \e);
\draw[orange, dashed] (1/2 + \e, 1/2- \e, 0) -- (\e, 0,1- \e);
\node (point) at (1/8 + \e, 1/8 -\e/2, 3/4 - \e/2) {$\bullet$};
%\draw[orange, very thick] (0.3,0.7,0) -- (0.3,0.7,2);
%\draw[orange, very thick] (0.7,0.3,0) -- (0.7,0.3,2);
\end{scope}
\node[inner sep=2pt] (levsetmin1) at (1/2, 1/4,1/4) {$\Gamma_{-1}$};
\node[inner sep=2pt] (levset1) at (1/4,1/2,1/4) {$\Gamma_{1}\!\!$};
%\draw[-latex,thin] (levset2) edge[in=270,out=90] (0.7,0.3,0);
%\draw[-latex,thin] (levset1) edge[in=270,out=90] (0.3,0.7,0);
\placefiglabel{(Q2)}
\end{tikzpicture}
\end{figure}
%\end{tabular}\label{tab:quadrants-simplices}
%\end{figure}

\section{Breadth of framework: Discrete-valued predictions}\label{sec:discrete-predictions-refactored}
In discrete prediction settings (Quadrants 1 and 2), we can apply Lemma~\ref{lem:convex-flats-inf-dim} to observe lower bounds on convex prediction dimension that subsume previous bounds given by~\citet{ramaswamy2016convex}.
Their lower bound 
%(the only non-trivial one to our knowledge) 
is given by the co-dimension of the \emph{feasible subspace dimension} $\Sc_\C(p)$ of a convex set $\C$ at a given distribution $p$ such that $p \in \C$, which is loosely the highest dimension subspace of $\C$ containing a neighborhood around $p$.
The feasible subspace dimension uses ``local'' information of the property to construct lower bounds, while $d$-flats allow us to also use ``global'' information, while subsuming the previous results about local information.

\begin{proposition}\label{prop:discrete-bound-subsumes-ramaswamy}
	Suppose we are given a \jessie{elicitable?} property $\gamma: \simplex \toto \R$.
	Fix $p \in \simplex$ and take $r \in \R$ such that $p \in \gamma_r$.  
	If there is a $d$-flat $F$ such that $p \in F \subseteq \gamma_r$, then $\codim(\Sc_{\gamma_r}(p)) \leq d$.
\end{proposition}

In Appendix~\ref{sec:finite-calib}, we show examples where the $d$-flats bound both matches the feasible subspace dimension bound and strictly improves it in order to emphasize that our result is strictly broader.

\section{Strength of framework: Continuous-valued predictions}\label{sec:contin-consis}

@@

\bibliographystyle{plainnat}
\bibliography{diss,extra}

@@

\section{Previous lower bounds and comparisons}\label{sec:finite-calib}

\begin{figure}[htb]
	\centering
	\begin{tabular}{cc}
		%		\includegraphics[width=.35\textwidth]{tikz/fsd-bound.pdf} & 
		%		\includegraphics[width=.35\textwidth]{tikz/flats-bound.pdf} \\
	\end{tabular}
	\caption{(Left) Feasible subspace dimension $\Sc_{\gamma_\bot}(\bullet) = 2$ and $\Sc_{\gamma_\bot}({\color{blue}\star}) = 1$, giving the bound $\conscvx(\ell^{abs}) \geq 3- 1-1 = 1$.
		(Right) No $1$-flat through $\bullet$ (a line since $\bullet \in \relint{\simplex})$  stays fully contained in $\gamma_\bot$, so $\conscvx(\ell^{abs}) \geq 2$.\jessie{FIX FIGURES}}
	\label{fig:fsd-flats-abstain}
\end{figure}
